---
title: "Sensitivity on gender equality in Italian parliament during the XVII legislature"
subtitle: "Report capstone project 2022"
author: "Erica Ravarelli"
output: pdf_document
---

```{r setup, include=FALSE}
library(SPARQL)
library(here)
library(vroom)
library(stringr)
library(tidyverse)
library(lubridate)
library(tidytext)
library(tidyr)
library(tm)
library(dplyr)
library(tables)
library(knitr)
library(forcats)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 99)
geip17 <- vroom(here("Data.csv/geip17.csv"))
tot2015 <- vroom(here("Data.csv/tot2015.csv"))
```

# Research question

This short research is aimed at investigating the relation between gender and tendency to talk about gender equality in Italian parliament. An insight on the relation between party ideology and tendency to talk about women rights will be given, too. The analysis focuses on the XVII legislature, with a specific attention to 2015.

Previous research have pointed out the fact that political representation is influenced by personal identity traits of the politician, specifically by his/her gender and race. Evidences about the relation between belonging political party and tendency to pay attention to gender equality have been found, too, with a strong association between left parties and sensitivity on gender equality issues. However, there is a lack of studies exploring these relations in the Italian context. My research aims at filling this gap.\
The selection concerning time period has been driven by the fact that the highest number of women deputies has been reached in the context of the last legislature, if we exclude the current one which is still ongoing. Moreover, the year 2015 has been the only one in which no change of government took place, suggesting that parliamentary activity hasn't been altered by external interferences.

## Research structure and methods

This project is structured as follows: in the first section, I will briefly describe the data collection process, in the second one I will explain how text mining has been carried out, whereas in the last section I will ho through the descriptive part of the analysis and I will explain how I applied the chi-squared test.

# Technical overview

The whole R scripts are available in [this](https://github.com/DataAccess2020/Capstone_project) Github repository. Therefore, in this short report I will only include and comment the most important passages of the analysis and the related lines of code. However, this does not affect the accessibility and reproducibility of the workflow.

## R packages

In the following lines I list all the packages used in my research.

```{r, message=FALSE}
library(SPARQL)
library(here)
library(vroom)
library(stringr)
library(tidyverse)
library(lubridate)
library(tidytext)
library(tidyr)
library(tm)
library(dplyr)
library(tables)
library(forcats)
library(ggplot2)
```

## Data collection

Data have been collected through the [Open Data portal of Camera dei Deputati](https://dati.camera.it/it/), in particular the Virtuoso endpoint has been queried using the SPARQL language; the returned data have been imported into R thanks to the SPARQL package. In both queries, the scraped url is the following one

```{r, site }
url="http://dati.camera.it/sparql"
```

The first query allowed to import data for the whole XVII legislature, in particular I asked for discussion topic, personal details of the deputy, belonging parliamentary group and office beginning and ending dates, along with intervention date. These last three variables were used in order to apply a filter: indeed, specifying that I was interested only in interventions which had taken place between the starting and the ending office dates have been fundamental in order to avoid including unwanted results. The `intervento` and the `discussione` variables allowed to check that no duplicates were included. Finally, only discussions which took place during the XVII legislature have been selected.

```{r, first query, eval=F}
"SELECT DISTINCT 
?argomento ?nome ?cognome ?genere ?dataNascita 
?gruppo_parlamentare ?background ?inizioIncarico ?fineIncarico ?data_intervento 
?intervento ?discussione

WHERE { 
##define the variable 'dibattito' and select all those belonging to the XVII legislature. 
##define the variable 'discussione' starting from the variable 'dibattito'
	?dibattito a ocd:dibattito;
                 ocd:rif_leg <http://dati.camera.it/ocd/legislatura.rdf/repubblica_17>;
                 ocd:rif_discussione ?discussione.

##define the variable 'argomento' starting from 'discussione'. Ask for 'data_intervento',
##which will be useful to fix a bug 
##select only discussions which have the word 'donna' or 'donne' in their title
 
  ?discussione ocd:rif_intervento ?intervento;
               rdfs:label ?argomento.
  FILTER(regex(?argomento, 'donn(e|a)', 'i')).
  OPTIONAL{?discussione dc:date ?data_intervento.}
  
##define the variable 'deputato' (intervenuto) starting from the variable 'intervento'
##ask for name, surname, gender and profesional background of the deputy
##define the variable 'intervento' and 'background'
 
  ?intervento ocd:rif_deputato ?deputato.
  ?deputato foaf:firstName ?nome; foaf:surname ?cognome; foaf:gender ?genere.
  ?deputato ocd:rif_ufficioParlamentare ?ufficioparl; rdfs:label ?incarico.
  OPTIONAL {?deputato dc:description ?background.}
 
##define the variable 'gruppo_parlamentare' (passing through 'aderisce') 
  
	?deputato ocd:aderisce ?aderisce.
	?aderisce rdfs:label ?gruppo_parlamentare.
  
##define the variables 'inizioIncarico' and 'fineIncarico', which will be useful
to fix a bug

  OPTIONAL{?aderisce ocd:endDate ?fineIncarico.}
  OPTIONAL{?aderisce ocd:startDate ?inizioIncarico.}
  
##ask for 'data di nascita', starting from 'persona'
 
  ?persona foaf:surname ?cognome; foaf:firstName ?nome; foaf:gender ?genere.
  ?persona <http://purl.org/vocab/bio/0.1/Birth> ?nascita.
  ?nascita <http://purl.org/vocab/bio/0.1/date> ?dataNascita. 
  
##specify that I only want deputies who interveined during their mandate  
##(don't take the same deputy more than once only because they 
changed office/parliamentary group) 
 
  FILTER(?data_intervento >= ?inizioIncarico && ?data_intervento < ?fineIncarico).
 
}
ORDER BY ?cognome"
```

In the second query an additional filter was added, in order to select only debates which took place during 2015. Moreover, a loop has been needed because the Virtuoso endpoint allows to scrape no more than 10000 observations within the same call. Therefore, we firstly have to generate an empty tibble which will then be filled-in with scraped data. Each call allowed to scrape 5000 results, then system was set to sleep for 1 second in order to ask for data in a polite way.

```{r, second query, eval=F}

"SELECT DISTINCT ?argomento ?nome ?cognome ?genere ?gruppo_parlamentare
?background ?inizioIncarico ?fineIncarico ?data_intervento ?intervento ?discussione
WHERE  {
  {
    SELECT DISTINCT ?argomento ?nome ?cognome ?genere ?gruppo_parlamentare 
    ?background ?inizioIncarico ?fineIncarico ?data_intervento ?intervento ?discussione
                      
                     WHERE  {
##define the variable 'dibattito' and select all those belonging to the XVII legislature. 
##define the variable 'discussione' starting from the variable 'dibattito'
	?dibattito a ocd:dibattito;
               ocd:rif_leg <http://dati.camera.it/ocd/legislatura.rdf/repubblica_17>.
  ?dibattito ocd:rif_discussione ?discussione.

##define the variable 'argomento' starting from 'discussione'. Ask for 'data_intervento',
##which will be useful to fix a bug 
 
  ?discussione ocd:rif_intervento ?intervento.
  ?discussione rdfs:label ?argomento.
  ?discussione dc:date ?data_intervento.
  
##define the variable 'deputato' (intervenuto) starting from the variable 'intervento'
##ask for name, surname, gender and professional background of the deputy
##define the variable 'intervento' and 'background'
 
  ?intervento ocd:rif_deputato ?deputato.
  ?deputato foaf:firstName ?nome; foaf:surname ?cognome; foaf:gender ?genere.
  OPTIONAL {?deputato dc:description ?background.}
 
  
##define the variable 'gruppo_parlamentare' (passing through 'aderisce') 
  
	?deputato ocd:aderisce ?aderisce.
	?aderisce rdfs:label ?gruppo_parlamentare.
  
##define the variables 'inizioIncarico' and 'fineIncarico', 
which will be useful to fix a bug

  OPTIONAL{?aderisce ocd:endDate ?fineIncarico.}
  OPTIONAL{?aderisce ocd:startDate ?inizioIncarico.}
  
##specify that I only want deputies who interveined during their mandate  
##(don't take the same deputy more than once only because 
##they changed office/parliamentary group) 
 
  FILTER(?data_intervento >= ?inizioIncarico && ?data_intervento < ?fineIncarico).
  FILTER(REGEX(?data_intervento,'^2015','i')).
}
ORDER BY ?cognome}
  }
LIMIT 10000
OFFSET"

tot2015 <- tibble()

#offset to ask for more than 10 000 results------ 
query_offset <- c("0","5000","10000", "20000", "30000", "40000", 
                  "50000", "60000")


i <- 0
for (i in 1:length(query_offset)) {
  interventi_deputati <- str_c(query,
                               query_offset[i],
                               sep = " ")
  final_dataset <- SPARQL(url, interventi_deputati)
  
  tot2015 <- rbind(tot2015, final_dataset$results)
  Sys.sleep(1)
}
```

After collecting them, data were saved in a csv format through the `write.csv()` command, in order to avoid having to query the endpoint more than needed.

```{r, save, eval=F}
write.csv(geip17, here::here("Data.csv/geip17.csv"))
write.csv(tot2015, here::here("Data.csv/tot2015.csv"))
```

## Data cleaning

Once collected and stored, data have been cleaned with the aim of preparing the analysis.

### XVII legislature

First of all, **dates** have been converted into a readable format through the `lubridate` package.

```{r, dates}
geip17$dataNascita <- ymd(geip17$dataNascita)
geip17$data_intervento <- ymd(geip17$data_intervento)
geip17$inizioIncarico <- ymd(geip17$inizioIncarico)
geip17$fineIncarico <- ymd(geip17$fineIncarico)
```

After that, regular expressions were used in order to recode the variable concerning the deputies' belonging **parliamentary group**. In particular:

-   The Five Star Movement has been considered as a stand-alone party, taking as valid the narration about not being neither a left nor a right part which was loudly upheld by the Movement members.

-   Analogously, but for a different and quite self-explanatory reason, deputies belonging to the mixed group were not assigned to a different category.

-   The `left` category includes: "movimento democratico e progressista", "partito democratico", "sinistra ecologia libertà", "sinistra italiana".

-   The `centre` category includes: "alternativa popolare-centristi per l'europa, per l'Italia".

-   The `right` category includes: "forza italia-il popolo della libertà-Berlusconi presidente", "fratelli d'Italia-alleanza nazionale", "il popolo della libertà-Berlusconi presidente", "lega nord e autonomie", "nuovo centrodestra".

```{r, regex gr parl}
m5s <- str_detect(geip17$gruppo_parlamentare, "STELLE")

mis <- str_detect(geip17$gruppo_parlamentare, "MISTO") 

sx <- str_detect(geip17$gruppo_parlamentare, "SINISTRA|DEMOCRATICO")

cen <- str_detect(geip17$gruppo_parlamentare, "CIVIC|POPOLARE|PER")

dx <- str_detect(geip17$gruppo_parlamentare, "POPOLO|FRATELLI|LEGA|DESTRA")
```

After making these decisions in terms of ideological positioning of parliamentary groups, the variable `gruppo_parl` has been generated through an `ifelse` command.

```{r, gruppo_parl}
gruppo_parl <- ifelse(m5s == TRUE,"M5S",
                      ifelse(mis == TRUE, "Misto",
                             ifelse(sx == TRUE, "Sinistra",
                                    ifelse(cen==TRUE, "Centro",
                                           ifelse(dx==TRUE, "Destra",
                                                  "altro")))))

geip17$gruppo_parlamentare <- gruppo_parl
```

Finally, I found out that in some cases deputies didn't actually talk about women: these observations had been included in my data set because the word "donna" was part of a surname ("Caradonna", "di Donna") or because deputies were actually talking about Jesus' mother, so the matched pattern was actually "Madonna". I, therefore, used the `str_match_all` command, which returns a matrix with a column for each observation pattern. I converted this matrix into the `to_remove` tibble, then I filtered the original data set, removing all observations in which the variable `x2` in `to_remove` was equal to `charachter(0)` (meaning that it did not match the pattern I was interested in removing).

```{r, remove obs}
to_remove <- str_match_all(geip17$argomento, 
                           "Renato Di Donna|Salvatore Caradonna|Madonna")

to_remove <- tibble(
  x1=1:366,
  x2=to_remove)

geip17 <- filter(
  geip17,
  to_remove$x2 %in% "character(0)")
```

### Year 2015

```{r, the same as above, include=FALSE, }
tot2015$data_intervento <- ymd(tot2015$data_intervento)
tot2015$inizioIncarico <- ymd(tot2015$inizioIncarico)
tot2015$fineIncarico <- ymd(tot2015$fineIncarico)

m5s <- str_detect(tot2015$gruppo_parlamentare, "STELLE")
mis <- str_detect(tot2015$gruppo_parlamentare, "MISTO")
sx <- str_detect(tot2015$gruppo_parlamentare, "SINISTRA|DEMOCRATICO")
cen <- str_detect(tot2015$gruppo_parlamentare, "CIVIC|POPOLARE|PER")
dx <- str_detect(tot2015$gruppo_parlamentare, "POPOLO|FRATELLI|LEGA|DESTRA")

gruppo_parl <- ifelse(m5s == TRUE,"M5S",
                      ifelse(mis == TRUE, "Misto",
                             ifelse(sx == TRUE, "Sinistra",
                                    ifelse(cen==TRUE, "Centro",
                                           ifelse(dx==TRUE, "Destra",
                                                  "altro")))))

tot2015$gruppo_parlamentare <- gruppo_parl
```

For what concerns the dataset with all parliamentary interventions of 2015, I went through the same operations as those explained above for dates and parliamentary group, then I recoded the variable `argomento` because I needed it to become a dummy: the deputy told about women/(s)he didn't.

```{r, argomento 2015}
match <- str_match_all(tot2015$argomento, "donn(e|a)")

tibble_match <- tibble(
  x1=1:58086,
  x2=match
)
tot2015$argomento <- ifelse(match %in% "character(0)", "any topic", 
                            "women rights/gender equality")
```

## Text mining XVII legislature

After carrying out a bit of data cleaning, I aimed at finding out within which scope deputies had been talking about women during the XVII legislature. In practical terms, this meant examining which words tended to immediately follow the word "donna". This task has been carried out through the `tidytext` package. First of all, I removed numbers from the variable `argomento`, then I used the `tm` package to download a list of Italian stop words, which I turned into the tibble `manualswords`. In this tibble I included also other words I didn't want to consider in my analysis.

```{r, stop words}

geip17$argomento <- removeNumbers(geip17$argomento)

italiansw <- tm::stopwords("italian")

manual_swords <-  c("d", "n", "s", "nonch")
comb_mswords <- c(manual_swords, italiansw)

manualswords <- tibble(
  x1=comb_mswords
)

```

After that, I split the variable `argomento` into consecutive sequences of words (n-grams), including the `token = "ngrams"` argument. I set `n=2` in order to obtain bigrams, so that I could examine pairs of two consecutive words. I, therefore, obtained the variable `split_argomento`.

```{r, bigrams}
bigrams <- geip17 %>% 
  select(argomento, genere) %>% 
  unnest_tokens(split_argomento, argomento, token="ngrams", n=2)
```

To remove stop words from this variable I had to split the column into its single words, and to filter the newly created tibble in order to remove words belonging to the `manualswords` tibble.

```{r, warning = FALSE}
bigrams_sep <- bigrams %>% 
  separate(split_argomento, c("parola1", "parola2"), sep = " ")

filterbigr <- bigrams_sep %>% 
  filter(!parola1 %in% manualswords$x1) %>% 
  filter(!parola2 %in% manualswords$x1)
```

At this point I could recombine these single words through the `unite` function, exclude bigrams in which the word "donne" was absent through the `grepl` function, and finally examine the most common bigrams (not containing stop words) using the `count()` function.

```{r, bigrams count}

bigramsunited <- filterbigr %>%
  unite(noswords, parola1, parola2, sep = " ")

greplbu <- bigramsunited %>% 
  filter(grepl(pattern="donn[ea]", noswords, ignore.case = TRUE))

bigrams_count <- greplbu %>%
  count(noswords)

```

The outcome of this analysis is reported in *Figure 1*.

```{r, out.width = '78%', fig.align = 'center', fig.cap = 'Most common bigrams containing the word "donna/e", XVII legislature', message = F, echo = F}
bigrams_count %>%
  mutate(noswords = fct_reorder(noswords, n)) %>%
  ggplot(aes(x=noswords, y=n)) +
  geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
  coord_flip() +
  xlab("") +
  theme_bw()
```

This result suggests that deputies told about women within "auditions", so during interventions which are meant to allow commissions to deepen their knowledge about a specific topic. This first result is, therefore, not very useful to understand the precise scope of speeches mentioning women. However, we can say that during the XVII legislature deputies told about women within the scope of pensions ("opzione donna"), which confirms an historical focus of Italian welfare state, and, as a general rule, when violence episodes took place or in order to address the issue of violence against women.

## Data viasualisation

At this point, I explored the relation between gender and frequency of interventions about women rights, controlling for ideological placement of the deputy. This was done through a bar chart in which the proportion of women and man deputies who told about women in their interventions was visualised within their belonging ideological position (*Figure 2*). The outcome suggests that left wing women told about gender equality more than women from other ideological groups. Moreover, the hypothesis according to which women are more likely to talk about women rights in parliament is only confirmed for what concerns left women.

```{r, out.width = '80%', fig.align = 'center', fig.cap = 'How much men and women deputies told about gender equality during the XVII legislature', message = F, echo = F}

propt17 <- prop.table(table(geip17$genere, geip17$gruppo_parlamentare), 2)
propt17 <- as.data.frame(propt17)

propt17 %>%  
  ggplot(aes(x=Var2, y=Freq, fill = factor(Var1))) +
  geom_bar(stat="identity", position = "fill", alpha=.6, width=.4) +
  labs(x = "parliamentary group",
       y = "frequency",
       fill="gender") +
       theme_bw()
```

The dataset with all interventions which took place in 2015 allowed to distinguish between two categories of discussions: those in which deputies mentioned the word "women" and those in which they didn't. In order to add this distinction, I firstly generated a three way table using the `xtabs` command, then I added margins in a way that allowed me to set the sum of women and men deputies to 100 for each category of the variable `argomento`. After converting the table into a tibble and removing sums, I could use it in my bar chart (*Figure 3*).

```{r, tables 2015}
#three way table
trevtable <- xtabs(~argomento+genere+gruppo_parlamentare, data=tot2015)
dimnames(trevtable) <- list(argomento = c("any topic", "gender eq/w. rights"),
                            genere = c("women", "men"),
                            gruppo_parlamentare=c("centre", "right", "M5S", 
                                                  "mix", "left"))
#male+female=100 for both categories of 'argomento'
final_table <- ftable(addmargins(prop.table(trevtable, c(1,3)), 2))*100
#convert the table into a tibble and take off the raws in which group = sum
percentages <- as_tibble(final_table)
percentages <- filter(
  percentages,
  !Freq == 100.00000)
```

The second bar chart clearly shows that the proportion of women deputies who intervened is much higher if the discussion is about gender equality. This is particularly true if we consider right and left women, whereas for other ideological positions the proportion of intervened men is higher than 0.5 even if the discussion is about gender equality.

```{r, out.width = '90%', fig.align = 'center', fig.cap = 'How much men and women deputies told about gender equality in 2015', message = F, echo = F}

percentages %>%
  ggplot(aes(x=gruppo_parlamentare, y=Freq, fill = factor(genere))) +
  geom_bar(stat="identity", position = "fill", alpha=.6, width=.4) +
  facet_wrap(~argomento, scale="free") +
  labs(x = "parliamentary group",
       y = "frequency",
       fill="gender") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title.align=0.5)

```

Finally, I calculated the chi-square statistic which allowed to detect the existence of an association between the variables `genere` and `gruppo parlamentare` and the variable `argomento`. The value of the chi-square is very high, but this could be due to the large sample size. I, therefore, analysed the standardized residuals, to find out when the association is positive and when it is negative. Results are reported in the table below and they confirm the hypothesis according to which left women are slightly more likely to talk about gender equality than women from other ideological positions, with the exception of women from the 'mix' group, because the observed frequency exceeds the one we would have observed if there was independence. However, if we consider left women, the observed frequency exceeds the expected one also for the other category of the variable `argomento`.

```{r, chisq,  message = F, echo = F, warning=F}
trevtable <- ftable(trevtable)
chisq <- chisq.test(trevtable)
tibblechisq <- tibble(
  Chisquare=chisq[["statistic"]],
  df=chisq[["parameter"]],
  pvalue=chisq[["p.value"]]
)
kable(tibblechisq)
```

```{r, fig.cap = 'standardixed residuals', message = F, echo = F}
tab <- round(chisq$residuals, 3)
tab
```
## Limits of the analysis and conclusion

The first limit of this analysis is due my lack of expertise in using the SPARQL language, which caused the decision of considering only discussions which took place in 2015 rather than selecting the whole period corresponding to Renzi govern (from February 2014 to December 2016).

Secondly, this analysis relies on a strong assumption according to which mentioning the word "donne" means talking about women rights and/or gender equality. It comes without saying that this is not always true. To try reducing the possibility of including intervention which were not relevant for the analysis, I decided to carry out text mining. However, the most common bigram turned out to be 'donne audizione', which, as already mentioned, does not provide the anylisis with a useful information about the precise scope within which the word "donne" was mentioned. 

The third limit of this short research consists of the absence of a quantitative prediction in terms of how much being a women determines being more likely to talk about issues concerning gender equality. 

Having said that, this research allows to point out the existence of differences in terms of parliamentary interventions about women rights, since the proportion of men who talk about gender equality issues is lower than the one of men who talk about other topics for all ideological positions. However, the hypothesis according to which the proportion of women who talk about gender equality is higher than the one of men is confirmed only for women belonging to right and left parties.